{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}

### 2016-08-30 ident_manu.tex
updating figures tables

\begin{singlespace}
\title{{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^1$, Lisa L. Lowe$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:lehrter.john@epa.gov}{lehrter.john@epa.gov}}}
	\\\\{\textit {\normalsize $^2$Lockheed Martin IS \& GS - Civil, supporting the USEPA}}
	\\{\textit {\normalsize Resarch Triangle Park, NC 27709}}
	\\{\textit {\normalsize Phone: 919-541-3985,}}
	\\{\textit {\normalsize Email: \href{mailto:lowe.lisa@epa.gov}{lowe.lisa@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Wed Aug 24 11:39:30 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

Results from FishTank are based on time-dependent differential equations that describe energy flow between phytoplankton and zooplankton groups as affected by nutrient uptake rates, organic matter inputs and losses, inherent optical properties, sediment diagenesis, and temperature (\citealt{Penta08,Eldridge10}, see appendix in \citetalias{LehrterIR}).  A total of 108 equations are estimated at each time step to return a value for each of the 36 state variables described by the model.  In addition to the initial conditions, 251 parameter values for each of the equations are also supplied at model execution.  These parameters define relationships among fixed effects in the equations and represent ecological properties described by the model that influence hypoxia formation.  Values for each of the parameters were based on estimates from the literature, field or laboratory-based measurements, or expert knowledge in absence of the former.  As such, a sensitivity analysis of parameter values is warranted given that, for example, literature or field-based estimates may not apply under all scenarios or expert knowledge is not completely certain \citep{Refsgaard07}.  The sensitivity of \ac{do} to perturbations of all relevant parameters for the 108 equations was estimated from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006 by running FishTank at a timestep of five minutes.

Irrelevant parameters were removed for several reasons from the analysis.  Parameters that were removed were not relevant for the \ac{zerod} model (i.e., hydrodynamic parameters), were considered physical constants, or had no effect given initial conditions.  Additionally, FishTank includes six phytoplankton and two zooplankton groups to describe complexity in community structure and foodweb dynamics.  However, structural equations for each group do not vary considerably such that variation in parameter values primarily controls differences between the groups, e.g., large-bodied or small-bodied plankton, slow-growing or fast-growing plankton, etc.  Initial analyses indicated that parameter sensitivity of dissolved oxygen was identical within the six phytoplankton and zooplankton groups.  To remove obvious redundancies in the model, the sensitivity analyses were conducted using only one phytoplankton and one zooplankton group.  The final parameter set that was evaluated included 51 parameters that were further grouped into one of six categories based on applicable biogeochemical components of the model: optics ($n = $ 4 parameters), organic matter (12), phytoplankton (22), temperature (2), and zooplankton (11).  A full description of the model parameters is available as an appendix in \citetalias{LehrterIR}.

@@

%latex.default(totab, file = "", rowlabel = "Description", caption = cap.val,     caption.loc = "top", rowname = Description, rgroup = unique(cats),     n.rgroup = as.numeric(table(cats)), size = tabsize, label = paste0("tab:",         tablab))%
\begin{table}[!tbp]
{\footnotesize
\caption{Sensitivities of \ac{do} to perturbation of indivividual parameters.  Sensitivities are based on a 50\% increase from the initial parameter value, where $L1$ summarizes differences in model output from the default (see \cref{l1}).  Parameters that did not affect \ac{do} are not shown.  Parameters are grouped by categories as optics, temperature, phytoplankton, zooplankton, and organic matter.\label{tab:dosens}} 
\begin{center}
\begin{tabular}{llll}
\hline\hline
\multicolumn{1}{l}{Description}&\multicolumn{1}{c}{Parameter}&\multicolumn{1}{c}{Value}&\multicolumn{1}{c}{L1}\tabularnewline
\hline
{\bfseries Optics}&&&\tabularnewline
~~OMA specific absorption at 490 nm&\textit{astarOMA}&$0.1$&$0.02$\tabularnewline
~~Chla specific absorption at 490 nm&\textit{astar490}&$0.04$&$0.02$\tabularnewline
~~OMZ specific absorption at 490 nm&\textit{astarOMZ}&$0.1$&$5.94\times 10^{-5}$\tabularnewline
\hline
{\bfseries Temperature}&&&\tabularnewline
~~Optimum temperature for growth(C)&\textit{Tref(nospA+nospZ)}&$22$&$0.92$\tabularnewline
\hline
{\bfseries Phytoplankton}&&&\tabularnewline
~~initial slope of the photosynthesis-irradiance relationship&\textit{alpha}&$8.42\times 10^{-17}$&$102.96$\tabularnewline
~~maximum growth rate&\textit{umax}&$0.41$&$18.21$\tabularnewline
~~mortality coefficient&\textit{mA}&$0.1$&$1.62$\tabularnewline
~~phytoplankton basal respiration coefficient&\textit{respb}&$0.02$&$1.2$\tabularnewline
~~phytoplankton carbon/cell&\textit{Qc}&$1.35\times 10^{-6}$&$0.95$\tabularnewline
~~phytoplankton growth respiration coefficient&\textit{respg}&$0.1$&$0.35$\tabularnewline
~~minimum N cell-quota&\textit{QminN}&$6.08\times 10^{-9}$&$0.16$\tabularnewline
~~N-uptake rate measured at umax&\textit{vmaxN}&$4.1\times 10^{-8}$&$0.1$\tabularnewline
~~coefficient for non-limiting nutrient&\textit{aN}&$1$&$0.09$\tabularnewline
~~half-saturation constant for P&\textit{Kp}&$2.86$&$0.09$\tabularnewline
~~half-saturation constant for N&\textit{Kn}&$4.51$&$8.86\times 10^{-3}$\tabularnewline
~~P-uptake rate measured at umax&\textit{vmaxP}&$2.68\times 10^{-8}$&$3.49\times 10^{-3}$\tabularnewline
~~minimum P cell-quota&\textit{QminP}&$6.19\times 10^{-10}$&$1.09\times 10^{-3}$\tabularnewline
\hline
{\bfseries Zooplankton}&&&\tabularnewline
~~zooplankton carbon/individual&\textit{ZQc}&$3.13\times 10^{-4}$&$1.87\times 10^{-3}$\tabularnewline
~~Zooplankton mortality constant for quadratic mortality&\textit{Zm}&$7.2\times 10^{-4}$&$1.1\times 10^{-3}$\tabularnewline
~~zooplankton phosphorus/individual&\textit{ZQp}&$3.77\times 10^{-6}$&$1.72\times 10^{-4}$\tabularnewline
~~Zooplankton biomass-dependent respiration factor&\textit{Zrespb}&$0.1$&$1.21\times 10^{-4}$\tabularnewline
~~zooplankton nitrogen/individual&\textit{ZQn}&$6.95\times 10^{-5}$&$2.19\times 10^{-5}$\tabularnewline
\hline
{\bfseries Organic Matter}&&&\tabularnewline
~~turnover rate for OM1A and OM1G&\textit{KG1}&$50$&$0.9$\tabularnewline
~~decay rate of CDOM, 1/day&\textit{KGcdom}&$0.01$&$0.53$\tabularnewline
~~half-saturation concentration for O2 utilization&\textit{KO2}&$10$&$0.15$\tabularnewline
~~O2 concentration that inhibits denitrification&\textit{KstarO2}&$10$&$0.11$\tabularnewline
~~half-saturation concentration for NO3 used in denitrification&\textit{KNO3}&$10$&$0.02$\tabularnewline
~~turnover rate for OM2A and OM2G&\textit{KG2}&$50$&$5.27\times 10^{-4}$\tabularnewline
~~maximum rate of nitrification per day&\textit{nitmax}&$0.52$&$3.67\times 10^{-6}$\tabularnewline
~~NH4 rate constant for nitrification&\textit{KNH4}&$1$&$1.75\times 10^{-6}$\tabularnewline
\hline
\end{tabular}\end{center}}

@@

% sensitivity example of top

% identifiability boxplots

% parameter exclusion temp

% identifiability boxplot, p1, z1 only

% heuristic plots o2

% heuristic plots, all state

@@

\caption{Sensitivity values (L1, \cref{l1}) for local analyses of all state variables. Parameters are grouped by category: optics, organic matter, phytoplankton, zooplankton, temperature, and zoplankton.  See \cref{tab:dosens} for L1 values for \ac{do}.}\label{fig:sensalltile}
\end{figure}

>>>>

\begin{singlespace}
\title{{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^1$, Lisa L. Lowe$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:lehrter.john@epa.gov}{lehrter.john@epa.gov}}}
	\\\\{\textit {\normalsize $^2$Lockheed Martin IS \& GS - Civil, supporting the USEPA}}
	\\{\textit {\normalsize Resarch Triangle Park, NC 27709}}
	\\{\textit {\normalsize Phone: 919-541-3985,}}
	\\{\textit {\normalsize Email: \href{mailto:lowe.lisa@epa.gov}{lowe.lisa@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Wed Aug 24 11:58:31 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

Results from FishTank are based on time-dependent differential equations that describe energy flow between phytoplankton and zooplankton groups as affected by nutrient uptake rates, organic matter inputs and losses, inherent optical properties, sediment diagenesis, and temperature (\citealt{Penta08,Eldridge10}, see appendix in \citetalias{LehrterIR}).  A total of 108 equations are estimated at each time step to return a value for each of the 36 state variables described by the model.  In addition to the initial conditions, 251 parameter values for each of the equations are also supplied at model execution.  These parameters define relationships among fixed effects in the equations and represent ecological properties described by the model that influence hypoxia formation.  Values for each of the parameters were based on estimates from the literature, field or laboratory-based measurements, or expert knowledge in absence of the former.  As such, a sensitivity analysis of parameter values is warranted given that, for example, literature or field-based estimates may not apply under all scenarios or expert knowledge is not completely certain \citep{Refsgaard07}.

The sensitivity of \ac{do} to perturbations of all relevant parameters for the 108 equations was estimated from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006 by running FishTank at a timestep of five minutes. Irrelevant parameters were removed for several reasons; parameters were not relevant for the \ac{zerod} model (i.e., hydrodynamic parameters), were considered physical constants, or had no effect given initial conditions.  Additionally, FishTank includes six phytoplankton and two zooplankton groups to describe complexity in community structure and foodweb dynamics.  However, structural equations for each group are identical such that chosen parameter values primarily control differences between the groups, e.g., large-bodied or small-bodied plankton, slow-growing or fast-growing plankton, etc.  Initial analyses indicated that parameter sensitivity of dissolved oxygen was identical within the six phytoplankton and zooplankton groups.  To remove obvious redundancies in the model, the sensitivity analyses were conducted using only one phytoplankton and one zooplankton group.  The final parameter set that was evaluated included 51 parameters that were further grouped into one of six categories based on applicable biogeochemical components of the model: optics ($n = $ 4 parameters), organic matter (12), phytoplankton (22), temperature (2), and zooplankton (11).  A full description of the model parameters is available as an appendix in \citetalias{LehrterIR}.

@@

% do sensitivity all categories
%latex.default(totab, file = "", rowlabel = "Description", caption = cap.val,     caption.loc = "top", rowname = Description, rgroup = unique(cats),     n.rgroup = as.numeric(table(cats)), size = tabsize, label = paste0("tab:",         tablab))%
\begin{table}[!tbp]
{\footnotesize
\caption{Sensitivity of \ac{do} to perturbations of indivividual parameters.  Sensitivities are based on a 50\% increase from the initial parameter value, where $L1$ summarizes differences in model output from the default (see \cref{l1}).  Parameters that did not affect \ac{do} are not shown.  Parameters are grouped by categories as optics, temperature, phytoplankton, zooplankton, and organic matter.\label{tab:dosens}} 
\begin{center}
\begin{tabular}{llll}
\hline\hline
\multicolumn{1}{l}{Description}&\multicolumn{1}{c}{Parameter}&\multicolumn{1}{c}{Value}&\multicolumn{1}{c}{L1}\tabularnewline
\hline
{\bfseries Optics}&&&\tabularnewline
~~Chla specific absorption at 490 nm&\textit{astar490}&$0.04$&$6.27\times 10^{-4}$\tabularnewline
~~OMA specific absorption at 490 nm&\textit{astarOMA}&$0.1$&$3.22\times 10^{-4}$\tabularnewline
~~OMZ specific absorption at 490 nm&\textit{astarOMZ}&$0.1$&$5.35\times 10^{-6}$\tabularnewline
\hline
{\bfseries Temperature}&&&\tabularnewline
~~Optimum temperature for growth(C)&\textit{Tref(nospA+nospZ)}&$22$&$1.8\times 10^{-4}$\tabularnewline
\hline
{\bfseries Phytoplankton}&&&\tabularnewline
~~phytoplankton carbon/cell&\textit{Qc}&$1.35\times 10^{-6}$&$0.71$\tabularnewline
~~initial slope of the photosynthesis-irradiance relationship&\textit{alpha}&$8.42\times 10^{-17}$&$0.03$\tabularnewline
~~mortality coefficient&\textit{mA}&$0.1$&$0.03$\tabularnewline
~~N-uptake rate measured at umax&\textit{vmaxN}&$4.1\times 10^{-8}$&$0.03$\tabularnewline
~~coefficient for non-limiting nutrient&\textit{aN}&$1$&$0.01$\tabularnewline
~~phytoplankton basal respiration coefficient&\textit{respb}&$0.02$&$0.01$\tabularnewline
~~half-saturation constant for P&\textit{Kp}&$2.86$&$0.01$\tabularnewline
~~minimum N cell-quota&\textit{QminN}&$6.08\times 10^{-9}$&$7.44\times 10^{-3}$\tabularnewline
~~maximum growth rate&\textit{umax}&$0.41$&$5.34\times 10^{-3}$\tabularnewline
~~half-saturation constant for N&\textit{Kn}&$4.51$&$1.56\times 10^{-3}$\tabularnewline
~~P-uptake rate measured at umax&\textit{vmaxP}&$2.68\times 10^{-8}$&$1.3\times 10^{-3}$\tabularnewline
~~minimum P cell-quota&\textit{QminP}&$6.19\times 10^{-10}$&$1.59\times 10^{-4}$\tabularnewline
~~phytoplankton growth respiration coefficient&\textit{respg}&$0.1$&$7.63\times 10^{-5}$\tabularnewline
\hline
{\bfseries Zooplankton}&&&\tabularnewline
~~zooplankton carbon/individual&\textit{ZQc}&$3.13\times 10^{-4}$&$4.51\times 10^{-3}$\tabularnewline
~~zooplankton nitrogen/individual&\textit{ZQn}&$6.95\times 10^{-5}$&$1.97\times 10^{-3}$\tabularnewline
~~Zooplankton biomass-dependent respiration factor&\textit{Zrespb}&$0.1$&$1.47\times 10^{-4}$\tabularnewline
~~Zooplankton mortality constant for quadratic mortality&\textit{Zm}&$7.2\times 10^{-4}$&$4.61\times 10^{-5}$\tabularnewline
~~zooplankton phosphorus/individual&\textit{ZQp}&$3.77\times 10^{-6}$&$2.89\times 10^{-5}$\tabularnewline
\hline
{\bfseries Organic Matter}&&&\tabularnewline
~~O2 concentration that inhibits denitrification&\textit{KstarO2}&$10$&$0.05$\tabularnewline
~~turnover rate for OM1A and OM1G&\textit{KG1}&$50$&$0.01$\tabularnewline
~~turnover rate for OM2A and OM2G&\textit{KG2}&$50$&$7.53\times 10^{-3}$\tabularnewline
~~half-saturation concentration for O2 utilization&\textit{KO2}&$10$&$6.49\times 10^{-3}$\tabularnewline
~~half-saturation concentration for NO3 used in denitrification&\textit{KNO3}&$10$&$6.36\times 10^{-3}$\tabularnewline
~~decay rate of CDOM, 1/day&\textit{KGcdom}&$0.01$&$8\times 10^{-4}$\tabularnewline
~~NH4 rate constant for nitrification&\textit{KNH4}&$1$&$2.9\times 10^{-4}$\tabularnewline
~~maximum rate of nitrification per day&\textit{nitmax}&$0.52$&$2.44\times 10^{-4}$\tabularnewline
\hline
\end{tabular}\end{center}}

@@

@@

\caption{Sensitivity values (L1, \cref{l1}) for local analyses of all state variables. Parameters are grouped by category: optics, organic matter, phytoplankton, zooplankton, temperature, and zoplankton.  See \cref{tab:dosens} for L1 values for \ac{do} and \cref{tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens} for the other state variables.}\label{fig:sensalltile}
\end{figure}

% identifiability boxplots
\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figs/identbox-1}

\caption{Identifiability (as $\gamma$, \cref{gameq}) of parameter subsets for \ac{do}.  Plots in (a) show identifiability by parameter categories and (b) shows identifiability by selecting the top 1 through 4 parameters in all categories.  Lines represent identifiability ranges for the possible combinations given the number of parameters in the set.  The temperature category is not shown because only \ac{do} was senstive to only one parameter.}\label{fig:identbox}
\end{figure}

% parameter exclusion temp
\begin{figure}[!ht]

{\centering \includegraphics[width=\textwidth]{figs/exclex-1}

\caption{Identifiability (as $\gamma$, \cref{gameq}) of organic matter parameters for subset combinations in \cref{fig:identbox}.  Identifiability is evaluated for subsets that excluded and included the parameters at the top of each plot. Identifiability of including all eight parameters is in \cref{fig:identbox}. Grey lines indicate potential thresholds at $\gamma = 10, 15$ for maximum acceptable identifiability.}\label{fig:exclex}
\end{figure}

% heuristic plots, all state
\begin{figure}[!ht]

{\centering \includegraphics[width=\textwidth]{figs/heurist_stts-1}

\caption[Identifiability (as ]{Identifiability (as $\gamma$, \cref{gameq}) of selecting parameters for all state variables. Parameters are selected by decreasing sensitivity independent of parameter categories. Grey lines indicate potential thresholds at $\gamma = 10, 15$ for maximum acceptable identifiability. Selection stops after $\gamma > 15$.}\label{fig:heurist_stts}
\end{figure}

% ammonium sensitivity all categories
%latex.default(totab, file = "", rowlabel = "Description", caption = cap.val,     caption.loc = "top", rowname = Description, rgroup = unique(cats),     n.rgroup = as.numeric(table(cats)), size = tabsize, label = paste0("tab:",         tablab))%
\begin{table}[!tbp]
{\footnotesize
\caption{Sensitivity of ammonium to perturbations of indivividual parameters.  Sensitivities are based on a 50\% increase from the initial parameter value, where $L1$ summarizes differences in model output from the default (see \cref{l1}).  Parameters that did not affect ammonium are not shown.  Parameters are grouped by categories as optics, temperature, phytoplankton, zooplankton, and organic matter.\label{tab:nh4sens}} 
\begin{center}
\begin{tabular}{llll}
\hline\hline
\multicolumn{1}{l}{Description}&\multicolumn{1}{c}{Parameter}&\multicolumn{1}{c}{Value}&\multicolumn{1}{c}{L1}\tabularnewline
\hline
{\bfseries Optics}&&&\tabularnewline
~~OMA specific absorption at 490 nm&\textit{astarOMA}&$0.1$&$0.01$\tabularnewline
~~Chla specific absorption at 490 nm&\textit{astar490}&$0.04$&$0.01$\tabularnewline
~~OMZ specific absorption at 490 nm&\textit{astarOMZ}&$0.1$&$4.77\times 10^{-5}$\tabularnewline
\hline
{\bfseries Temperature}&&&\tabularnewline
~~Optimum temperature for growth(C)&\textit{Tref(nospA+nospZ)}&$22$&$0.61$\tabularnewline
\hline
{\bfseries Phytoplankton}&&&\tabularnewline
~~initial slope of the photosynthesis-irradiance relationship&\textit{alpha}&$8.42\times 10^{-17}$&$43.95$\tabularnewline
~~maximum growth rate&\textit{umax}&$0.41$&$9.63$\tabularnewline
~~mortality coefficient&\textit{mA}&$0.1$&$1.32$\tabularnewline
~~phytoplankton basal respiration coefficient&\textit{respb}&$0.02$&$0.9$\tabularnewline
~~N-uptake rate measured at umax&\textit{vmaxN}&$4.1\times 10^{-8}$&$0.56$\tabularnewline
~~coefficient for non-limiting nutrient&\textit{aN}&$1$&$0.31$\tabularnewline
~~half-saturation constant for P&\textit{Kp}&$2.86$&$0.3$\tabularnewline
~~phytoplankton growth respiration coefficient&\textit{respg}&$0.1$&$0.21$\tabularnewline
~~minimum N cell-quota&\textit{QminN}&$6.08\times 10^{-9}$&$0.05$\tabularnewline
~~half-saturation constant for N&\textit{Kn}&$4.51$&$0.04$\tabularnewline
~~phytoplankton carbon/cell&\textit{Qc}&$1.35\times 10^{-6}$&$0.02$\tabularnewline
~~P-uptake rate measured at umax&\textit{vmaxP}&$2.68\times 10^{-8}$&$9.3\times 10^{-3}$\tabularnewline
~~minimum P cell-quota&\textit{QminP}&$6.19\times 10^{-10}$&$6.4\times 10^{-3}$\tabularnewline
\hline
{\bfseries Zooplankton}&&&\tabularnewline
~~zooplankton nitrogen/individual&\textit{ZQn}&$6.95\times 10^{-5}$&$0.18$\tabularnewline
~~Zooplankton biomass-dependent respiration factor&\textit{Zrespb}&$0.1$&$0.17$\tabularnewline
~~Zooplankton mortality constant for quadratic mortality&\textit{Zm}&$7.2\times 10^{-4}$&$1.41\times 10^{-3}$\tabularnewline
~~zooplankton phosphorus/individual&\textit{ZQp}&$3.77\times 10^{-6}$&$1\times 10^{-3}$\tabularnewline
~~zooplankton carbon/individual&\textit{ZQc}&$3.13\times 10^{-4}$&$1.85\times 10^{-4}$\tabularnewline
\hline
{\bfseries Organic Matter}&&&\tabularnewline
~~NH4 rate constant for nitrification&\textit{KNH4}&$1$&$1.1$\tabularnewline
~~maximum rate of nitrification per day&\textit{nitmax}&$0.52$&$0.71$\tabularnewline
~~decay rate of CDOM, 1/day&\textit{KGcdom}&$0.01$&$0.36$\tabularnewline
~~turnover rate for OM1A and OM1G&\textit{KG1}&$50$&$0.2$\tabularnewline
~~half-saturation concentration for O2 utilization&\textit{KO2}&$10$&$0.14$\tabularnewline
~~O2 concentration that inhibits denitrification&\textit{KstarO2}&$10$&$0.07$\tabularnewline
~~turnover rate for OM2A and OM2G&\textit{KG2}&$50$&$0.05$\tabularnewline
~~half-saturation concentration for NO3 used in denitrification&\textit{KNO3}&$10$&$0.01$\tabularnewline
\hline
\end{tabular}\end{center}}

% chlorophyll sensitivity all categories
%latex.default(totab, file = "", rowlabel = "Description", caption = cap.val,     caption.loc = "top", rowname = Description, rgroup = unique(cats),     n.rgroup = as.numeric(table(cats)), size = tabsize, label = paste0("tab:",         tablab))%
\begin{table}[!tbp]
{\footnotesize
\caption{Sensitivity of \ac{chla} to perturbations of indivividual parameters.  Sensitivities are based on a 50\% increase from the initial parameter value, where $L1$ summarizes differences in model output from the default (see \cref{l1}).  Parameters that did not affect \ac{chla} are not shown.  Parameters are grouped by categories as optics, temperature, phytoplankton, zooplankton, and organic matter.\label{tab:chlsens}} 
\begin{center}
\begin{tabular}{llll}
\hline\hline
\multicolumn{1}{l}{Description}&\multicolumn{1}{c}{Parameter}&\multicolumn{1}{c}{Value}&\multicolumn{1}{c}{L1}\tabularnewline
\hline
{\bfseries Optics}&&&\tabularnewline
~~OMA specific absorption at 490 nm&\textit{astarOMA}&$0.1$&$0.02$\tabularnewline
~~Chla specific absorption at 490 nm&\textit{astar490}&$0.04$&$0.02$\tabularnewline
~~OMZ specific absorption at 490 nm&\textit{astarOMZ}&$0.1$&$6.77\times 10^{-5}$\tabularnewline
\hline
{\bfseries Temperature}&&&\tabularnewline
~~Optimum temperature for growth(C)&\textit{Tref(nospA+nospZ)}&$22$&$0.96$\tabularnewline
\hline
{\bfseries Phytoplankton}&&&\tabularnewline
~~initial slope of the photosynthesis-irradiance relationship&\textit{alpha}&$8.42\times 10^{-17}$&$132.89$\tabularnewline
~~maximum growth rate&\textit{umax}&$0.41$&$20.34$\tabularnewline
~~mortality coefficient&\textit{mA}&$0.1$&$1.9$\tabularnewline
~~phytoplankton basal respiration coefficient&\textit{respb}&$0.02$&$1.33$\tabularnewline
~~phytoplankton carbon/cell&\textit{Qc}&$1.35\times 10^{-6}$&$0.91$\tabularnewline
~~phytoplankton growth respiration coefficient&\textit{respg}&$0.1$&$0.37$\tabularnewline
~~minimum N cell-quota&\textit{QminN}&$6.08\times 10^{-9}$&$4.69\times 10^{-3}$\tabularnewline
~~half-saturation constant for P&\textit{Kp}&$2.86$&$2.25\times 10^{-3}$\tabularnewline
~~coefficient for non-limiting nutrient&\textit{aN}&$1$&$2.13\times 10^{-3}$\tabularnewline
~~N-uptake rate measured at umax&\textit{vmaxN}&$4.1\times 10^{-8}$&$7.11\times 10^{-4}$\tabularnewline
~~half-saturation constant for N&\textit{Kn}&$4.51$&$2.05\times 10^{-4}$\tabularnewline
~~P-uptake rate measured at umax&\textit{vmaxP}&$2.68\times 10^{-8}$&$1.71\times 10^{-4}$\tabularnewline
~~minimum P cell-quota&\textit{QminP}&$6.19\times 10^{-10}$&$2.32\times 10^{-5}$\tabularnewline
\hline
{\bfseries Zooplankton}&&&\tabularnewline
~~zooplankton carbon/individual&\textit{ZQc}&$3.13\times 10^{-4}$&$6.79\times 10^{-5}$\tabularnewline
~~Zooplankton biomass-dependent respiration factor&\textit{Zrespb}&$0.1$&$1.46\times 10^{-5}$\tabularnewline
~~Zooplankton mortality constant for quadratic mortality&\textit{Zm}&$7.2\times 10^{-4}$&$1.07\times 10^{-5}$\tabularnewline
~~zooplankton phosphorus/individual&\textit{ZQp}&$3.77\times 10^{-6}$&$6.48\times 10^{-6}$\tabularnewline
\hline
{\bfseries Organic Matter}&&&\tabularnewline
~~decay rate of CDOM, 1/day&\textit{KGcdom}&$0.01$&$0.58$\tabularnewline
~~turnover rate for OM1A and OM1G&\textit{KG1}&$50$&$0.02$\tabularnewline
~~half-saturation concentration for O2 utilization&\textit{KO2}&$10$&$1.69\times 10^{-3}$\tabularnewline
~~O2 concentration that inhibits denitrification&\textit{KstarO2}&$10$&$1.38\times 10^{-3}$\tabularnewline
~~half-saturation concentration for NO3 used in denitrification&\textit{KNO3}&$10$&$1.96\times 10^{-4}$\tabularnewline
~~turnover rate for OM2A and OM2G&\textit{KG2}&$50$&$3.14\times 10^{-5}$\tabularnewline
\hline
\end{tabular}\end{center}}

% irradiance sensitivity all categories
%latex.default(totab, file = "", rowlabel = "Description", caption = cap.val,     caption.loc = "top", rowname = Description, rgroup = unique(cats),     n.rgroup = as.numeric(table(cats)), size = tabsize, label = paste0("tab:",         tablab))%
\begin{table}[!tbp]
{\footnotesize
\caption{Sensitivity of irradiance to perturbations of indivividual parameters.  Sensitivities are based on a 50\% increase from the initial parameter value, where $L1$ summarizes differences in model output from the default (see \cref{l1}).  Parameters that did not affect irradiance are not shown.  Parameters are grouped by categories as optics, temperature, phytoplankton, zooplankton, and organic matter.\label{tab:irrsens}} 
\begin{center}
\begin{tabular}{llll}
\hline\hline
\multicolumn{1}{l}{Description}&\multicolumn{1}{c}{Parameter}&\multicolumn{1}{c}{Value}&\multicolumn{1}{c}{L1}\tabularnewline
\hline
{\bfseries Optics}&&&\tabularnewline
~~Chla specific absorption at 490 nm&\textit{astar490}&$0.04$&$2.71\times 10^{-3}$\tabularnewline
~~OMA specific absorption at 490 nm&\textit{astarOMA}&$0.1$&$2.41\times 10^{-3}$\tabularnewline
~~OMZ specific absorption at 490 nm&\textit{astarOMZ}&$0.1$&$7.15\times 10^{-6}$\tabularnewline
\hline
{\bfseries Temperature}&&&\tabularnewline
~~Optimum temperature for growth(C)&\textit{Tref(nospA+nospZ)}&$22$&$2.29\times 10^{-4}$\tabularnewline
\hline
{\bfseries Phytoplankton}&&&\tabularnewline
~~phytoplankton carbon/cell&\textit{Qc}&$1.35\times 10^{-6}$&$8.09\times 10^{-3}$\tabularnewline
~~initial slope of the photosynthesis-irradiance relationship&\textit{alpha}&$8.42\times 10^{-17}$&$7.65\times 10^{-3}$\tabularnewline
~~mortality coefficient&\textit{mA}&$0.1$&$7.42\times 10^{-3}$\tabularnewline
~~phytoplankton basal respiration coefficient&\textit{respb}&$0.02$&$2.13\times 10^{-3}$\tabularnewline
~~maximum growth rate&\textit{umax}&$0.41$&$2.09\times 10^{-3}$\tabularnewline
~~minimum N cell-quota&\textit{QminN}&$6.08\times 10^{-9}$&$7.39\times 10^{-4}$\tabularnewline
~~N-uptake rate measured at umax&\textit{vmaxN}&$4.1\times 10^{-8}$&$5.74\times 10^{-4}$\tabularnewline
~~coefficient for non-limiting nutrient&\textit{aN}&$1$&$4.7\times 10^{-4}$\tabularnewline
~~half-saturation constant for P&\textit{Kp}&$2.86$&$4.44\times 10^{-4}$\tabularnewline
~~phytoplankton growth respiration coefficient&\textit{respg}&$0.1$&$6.64\times 10^{-5}$\tabularnewline
~~half-saturation constant for N&\textit{Kn}&$4.51$&$4.7\times 10^{-5}$\tabularnewline
~~P-uptake rate measured at umax&\textit{vmaxP}&$2.68\times 10^{-8}$&$3.48\times 10^{-5}$\tabularnewline
~~minimum P cell-quota&\textit{QminP}&$6.19\times 10^{-10}$&$2.45\times 10^{-6}$\tabularnewline
\hline
{\bfseries Zooplankton}&&&\tabularnewline
~~zooplankton carbon/individual&\textit{ZQc}&$3.13\times 10^{-4}$&$7.23\times 10^{-6}$\tabularnewline
~~Zooplankton biomass-dependent respiration factor&\textit{Zrespb}&$0.1$&$1.05\times 10^{-6}$\tabularnewline
~~Zooplankton mortality constant for quadratic mortality&\textit{Zm}&$7.2\times 10^{-4}$&$9.27\times 10^{-7}$\tabularnewline
~~zooplankton phosphorus/individual&\textit{ZQp}&$3.77\times 10^{-6}$&$4.47\times 10^{-7}$\tabularnewline
~~zooplankton nitrogen/individual&\textit{ZQn}&$6.95\times 10^{-5}$&$2.43\times 10^{-8}$\tabularnewline
\hline
{\bfseries Organic Matter}&&&\tabularnewline
~~decay rate of CDOM, 1/day&\textit{KGcdom}&$0.01$&$0.05$\tabularnewline
~~turnover rate for OM1A and OM1G&\textit{KG1}&$50$&$1.6\times 10^{-3}$\tabularnewline
~~half-saturation concentration for O2 utilization&\textit{KO2}&$10$&$1.61\times 10^{-4}$\tabularnewline
~~O2 concentration that inhibits denitrification&\textit{KstarO2}&$10$&$1.31\times 10^{-4}$\tabularnewline
~~half-saturation concentration for NO3 used in denitrification&\textit{KNO3}&$10$&$1.82\times 10^{-5}$\tabularnewline
~~turnover rate for OM2A and OM2G&\textit{KG2}&$50$&$2.95\times 10^{-6}$\tabularnewline
~~NH4 rate constant for nitrification&\textit{KNH4}&$1$&$7.41\times 10^{-9}$\tabularnewline
~~maximum rate of nitrification per day&\textit{nitmax}&$0.52$&$4.81\times 10^{-9}$\tabularnewline
\hline
\end{tabular}\end{center}}

% nitrate sensitivity all categories
%latex.default(totab, file = "", rowlabel = "Description", caption = cap.val,     caption.loc = "top", rowname = Description, rgroup = unique(cats),     n.rgroup = as.numeric(table(cats)), size = tabsize, label = paste0("tab:",         tablab))%
\begin{table}[!tbp]
{\footnotesize
\caption{Sensitivity of nitrate to perturbations of indivividual parameters.  Sensitivities are based on a 50\% increase from the initial parameter value, where $L1$ summarizes differences in model output from the default (see \cref{l1}).  Parameters that did not affect nitrate are not shown.  Parameters are grouped by categories as optics, temperature, phytoplankton, zooplankton, and organic matter.\label{tab:no3sens}} 
\begin{center}
\begin{tabular}{llll}
\hline\hline
\multicolumn{1}{l}{Description}&\multicolumn{1}{c}{Parameter}&\multicolumn{1}{c}{Value}&\multicolumn{1}{c}{L1}\tabularnewline
\hline
{\bfseries Optics}&&&\tabularnewline
~~Chla specific absorption at 490 nm&\textit{astar490}&$0.04$&$5.03\times 10^{-4}$\tabularnewline
~~OMA specific absorption at 490 nm&\textit{astarOMA}&$0.1$&$2.58\times 10^{-4}$\tabularnewline
